decompr: GVC decomposition in R

### 2015-02-09 draft.Rnw
recompile

The application of the Leontief insight to ICIOs as opposed to national Input-Output tables for our \textbf{Leontief} decomposition is straightforward.
\textbf{V} refers now to a vector of direct value added contribution of all industries across the different countries. 
Its dimension is correspondingly 1 x GN, where G is according to the number of countries.
\textbf{A} is now of dimension GN x GN and gives the industry flows including cross border relationships.
Since we are interested in the value added origins of exports we multiply these two matrices with a GN x GN matrix whose diagonal we fill with each industry's exports, $E$, 
such that the basic equation behind the source decomposition is given by $V(I-A)^{-1}E$ 
\footnote{When using the textbf{leontief_output} function, the value added multiplier is instead multiplied with each industry's output.}.
In a simple example with two countries (\textbf{k} and \textbf{l}) and industries (\textbf{i} and \textbf{j}) we can zoom in to see the matrices' content:

>>>>

The application of the Leontief insight to ICIOs as opposed to national Input-Output tables for our \textbf{Leontief} decomposition is straightforward.
\textbf{V} refers now to a vector of direct value added contribution of all industries across the different countries. 
Its dimension is correspondingly 1 x GN, where G is according to the number of countries.
\textbf{A} is now of dimension GN x GN and gives the industry flows including cross border relationships.
Since we are interested in the value added origins of exports we multiply these two matrices with a GN x GN matrix whose diagonal we fill with each industry's exports, $E$, 
such that the basic equation behind the source decomposition is given by $V(I-A)^{-1}E$ 
\footnote{When using the textbf{leontief\_output} function, the value added multiplier is instead multiplied with each industry's output.}
In a simple example with two countries (\textbf{k} and \textbf{l}) and industries (\textbf{i} and \textbf{j}) we can zoom in to see the matrices' content:

### 2015-02-09 draft.tex
recompile

\begin{equation}
VB = V (I-A)^{-1},
\end{equation}
where \textbf{V} is a N x N matrix with the diagonal representing the direct value added contribution of the industries,
\textbf{A} is the Input-Output coefficient matrix with dimension N x N, i.e. it gives the direct input flows between
industries required for 1\$ of output, $B = (I-A)^{-1}$ is then the so called Leontief inverse, and \textbf{T} indicates
a matrix transpose operation.
\textbf{VB} gives thus a N x N matrix of so called value added multipliers, which denote the amount of value added that the production of an industry's 1\$ of output or exports brings about in all other industries.
Looking from the perspective of the supplying industries, the matrix gives the value added that they contribute to the using industry's production.
If we multiply it with a N x N matrix whose diagonal specifies each industry's total output, we get value added origins as absolute values instead of shares.

The application of the Leontief insight to ICIOs as opposed to national Input-Output tables for our source decomposition
is straightforward.
\textbf{V} refers now to a vector of direct value added contribution of all industries across the different countries. Its dimension is correspondingly 1 x GN.
\textbf{A} is now of dimension GN x GN and gives the industry flows including cross border relationships.
Since we are interested in the value added origins of exports we multiply these two matrices with a GN x GN matrix whose diagonal we fill with each industry's exports, $E$, such that the basic equation behind the source decomposition is given by $V(I-A)^{-1}E$.
In a simple example with two countries (\textbf{k} and \textbf{l}) and industries (\textbf{i} and \textbf{j}) we can zoom in to see the matrices' content:

\begin{align}
\begin{split}
V (I- A &)^{-1} E=
\begin{pmatrix}
v_{k}^{i}& 0& 0& 0\\
0& v_{k}^{j}& 0& 0\\
0& 0& v_{l}^{i}& 0\\
0& 0& 0& v_{l}^{j}
\end{pmatrix}
*
\begin{pmatrix}
b_{kk}^{ii}& b_{kk}^{ij}& b_{kl}^{ii}& b_{kl}^{ij}\\
b_{kk}^{ji}& b_{kk}^{jj}& b_{kl}^{ji}& b_{kl}^{jj}\\
b_{lk}^{ii}& b_{lk}^{ij}& b_{ll}^{ii}& b_{ll}^{ij}\\
b_{lk}^{ji}& b_{lk}^{jj}& b_{ll}^{ji}& b_{ll}^{jj}
\end{pmatrix}
*
\begin{pmatrix}
e_{k}^{i}& 0& 0& 0\\
0& e_{k}^{j}& 0& 0\\
0& 0& e_{l}^{i}& 0\\
0& 0& 0& e_{l}^{j}
\end{pmatrix}
=
\\
&
\begin{pmatrix}
v_{k}^{i}b_{kk}^{ii}e_{k}^{i}& v_{k}^{i} b_{kk}^{ij}e_{k}^{j}& v_{k}^{i}b_{kl}^{ii}e_{l}^{i}& v_{k}^{i}b_{kl}^{ij}e_{l}^{j}\\
v_{k}^{j}b_{kk}^{ji}e_{k}^{i}& v_{k}^{j}b_{kk}^{jj}e_{k}^{j}& v_{k}^{j}b_{kl}^{ji}e_{l}^{i}& v_{k}^{j}b_{kl}^{jj}e_{l}^{j}\\
v_{l}^{i}b_{lk}^{ii}e_{k}^{i}& v_{l}^{i}b_{lk}^{ij}e_{k}^{j}& v_{l}^{i}b_{ll}^{ii}e_{l}^{i}& v_{l}^{i}b_{ll}^{ij}e_{l}^{j}\\
v_{l}^{j}b_{lk}^{ji}e_{k}^{i}& v_{l}^{j}b_{lk}^{jj}e_{k}^{j}& v_{l}^{j}b_{ll}^{ji}e_{l}^{i}& v_{l}^{j}b_{ll}^{jj}e_{l}^{j}
\end{pmatrix}
=
\begin{pmatrix}
vae_{kk}^{ii}& vae_{kk}^{ij}& vae_{kl}^{ii}& vae_{kl}^{ij}\\
vae_{kk}^{ji}& vae_{kk}^{jj}& vae_{kl}^{ji}& vae_{kl}^{jj}\\
vae_{lk}^{ii}& vae_{lk}^{ij}& vae_{ll}^{ii}& vae_{ll}^{ij}\\
vae_{lk}^{ji}& vae_{lk}^{jj}& vae_{ll}^{ji}& vae_{ll}^{jj}
\end{pmatrix}
\end{split}
\end{align}
\begin{align*}
&v_{c}^{s} = \frac{va_{c}^{s}}{y_{c}^{s}} = 1 - a_{kc}^{is} - a_{kc}^{js} - a_{lc}^{js} - a_{lc}^{is} \hspace{2cm} (c \in k,l \hspace{.5cm} s \in i,j),\\
&\begin{pmatrix}
b_{kk}^{ii}& b_{kk}^{ij}& b_{kl}^{ii}& b_{kl}^{ij}\\
b_{kk}^{ji}& b_{kk}^{jj}& b_{kl}^{ji}& b_{kl}^{jj}\\
b_{lk}^{ii}& b_{lk}^{ij}& b_{ll}^{ii}& b_{ll}^{ij}\\
b_{lk}^{ji}& b_{lk}^{jj}& b_{ll}^{ji}& b_{ll}^{jj}
\end{pmatrix}
=
\begin{pmatrix}
1-a_{kk}^{ii}& -a_{kk}^{ij}& -a_{kl}^{ii}& -a_{kl}^{ij}\\
-a_{kk}^{ji}& 1-a_{kk}^{jj}& -a_{kl}^{ji}& -a_{kl}^{jj}\\
-a_{lk}^{ii}& -a_{lk}^{ij}& 1-a_{ll}^{ii}& -a_{ll}^{ij}\\
-a_{lk}^{ji}& -a_{lk}^{jj}& -a_{ll}^{ji}& 1-a_{ll}^{jj}
\end{pmatrix}^{-1},
\end{align*}
and
\begin{equation*}
a_{cf}^{su} = \frac{inp_{cf}^{su}}{y_{f}^{u}}  \hspace{2cm} (c,f \in k,l \hspace{.5cm} s,u \in i,j),
\end{equation*}
where \(v_{s}^{c}\) gives the share of industry \emph{s}'s value added, \(va_{c}^{s},\) in output, \(y_{s}^{c}\), and \(e_{k}^{i}\) indicates gross exports. Finally, \(a_{su}^{cf}\) denotes the share of inputs, \(inp_{su}^{cf}\), in output.
The elements of the $V(I-A)^{-1}E$ or $vae$ matrix are our estimates for the country-industry level value added
origins of each country-industry's exports.
\textbf{decompr} implements this algorithm into R to automate the process of deriving the matrix. 
Equipped with it, researchers can calculate standard GVC indicators.
Examples include \citet{dahuetal01}'s Vertical Specialisation ratio at the industry-level, 
which can be derived by summing for each country and industry across the value added of all foreign countries and industries,
and \citet{rojoguno12}'s so-called VAX ratio.
Alternatively, the four dimensions of the matrix (source country, source industry, using country, using industry) allow for industry-level gravity-type estimations of value added trade flows.

@@

@@

We have designed the package using a modular structure, with an additional user interface function for increased ease of use.
The modulare structure enables users to break the computationally intensive analysis process down into several steps.
Furthermore, the modular structure enables user and other developers to build on top of basic data structures which are creaed by the \textbf{load\_tables\_vectors} function when implementing other decompositions or analyses.
However, a wrapper function (\textbf{decomp}) is also provided, which combines the use of the atomic functions into one.
Lastly, in addition to the \textbf{decompr} function, we provide a Rcmdr plugin \parencite{fox2005getting, fox2007extending}, which allows the user to perform the analytical process using a GUI (Grapical User Interface).

>>>>

\begin{equation}
VB = V (I-A)^{-1},
\end{equation}
where \textbf{V} is a N x N matrix with the diagonal representing the direct value added contribution of N industries,
\textbf{A} is the Input-Output coefficient matrix with dimension N x N, i.e. it gives the direct input flows between
industries required for 1\$ of output, and $B = (I-A)^{-1}$ is the so called Leontief inverse.
\textbf{VB} gives thus a N x N matrix of so called value added multipliers, which denote the amount of value added that the production of an industry's 1\$ of output or exports brings about in all other industries.
Looking from the perspective of the supplying industries, the matrix gives the value added that they contribute to the using industry's production.
If we multiply it with a N x N matrix whose diagonal specifies each industry's total output or exports, we get value added origins as absolute values instead of shares.

The application of the Leontief insight to ICIOs as opposed to national Input-Output tables for our \textbf{Leontief} decomposition is straightforward.
\textbf{V} refers now to a vector of direct value added contribution of all industries across the different countries. 
Its dimension is correspondingly 1 x GN, where G is according to the number of countries.
\textbf{A} is now of dimension GN x GN and gives the industry flows including cross border relationships.
Since we are interested in the value added origins of exports we multiply these two matrices with a GN x GN matrix whose diagonal we fill with each industry's exports, $E$, 
such that the basic equation behind the source decomposition is given by $V(I-A)^{-1}E$ 
\footnote{When using the textbf{leontief\_output} function, the value added multiplier is instead multiplied with each industry's output.}
In a simple example with two countries (\textbf{k} and \textbf{l}) and industries (\textbf{i} and \textbf{j}) we can zoom in to see the matrices' content:

\begin{align}
\begin{split}
V (I- A &)^{-1} E=
\begin{pmatrix}
v_{k}^{i}& 0& 0& 0\\
0& v_{k}^{j}& 0& 0\\
0& 0& v_{l}^{i}& 0\\
0& 0& 0& v_{l}^{j}
\end{pmatrix}
*
\begin{pmatrix}
b_{kk}^{ii}& b_{kk}^{ij}& b_{kl}^{ii}& b_{kl}^{ij}\\
b_{kk}^{ji}& b_{kk}^{jj}& b_{kl}^{ji}& b_{kl}^{jj}\\
b_{lk}^{ii}& b_{lk}^{ij}& b_{ll}^{ii}& b_{ll}^{ij}\\
b_{lk}^{ji}& b_{lk}^{jj}& b_{ll}^{ji}& b_{ll}^{jj}
\end{pmatrix}
*
\begin{pmatrix}
e_{k}^{i}& 0& 0& 0\\
0& e_{k}^{j}& 0& 0\\
0& 0& e_{l}^{i}& 0\\
0& 0& 0& e_{l}^{j}
\end{pmatrix}
=
\\
&
\begin{pmatrix}
v_{k}^{i}b_{kk}^{ii}e_{k}^{i}& v_{k}^{i} b_{kk}^{ij}e_{k}^{j}& v_{k}^{i}b_{kl}^{ii}e_{l}^{i}& v_{k}^{i}b_{kl}^{ij}e_{l}^{j}\\
v_{k}^{j}b_{kk}^{ji}e_{k}^{i}& v_{k}^{j}b_{kk}^{jj}e_{k}^{j}& v_{k}^{j}b_{kl}^{ji}e_{l}^{i}& v_{k}^{j}b_{kl}^{jj}e_{l}^{j}\\
v_{l}^{i}b_{lk}^{ii}e_{k}^{i}& v_{l}^{i}b_{lk}^{ij}e_{k}^{j}& v_{l}^{i}b_{ll}^{ii}e_{l}^{i}& v_{l}^{i}b_{ll}^{ij}e_{l}^{j}\\
v_{l}^{j}b_{lk}^{ji}e_{k}^{i}& v_{l}^{j}b_{lk}^{jj}e_{k}^{j}& v_{l}^{j}b_{ll}^{ji}e_{l}^{i}& v_{l}^{j}b_{ll}^{jj}e_{l}^{j}
\end{pmatrix}
=
\begin{pmatrix}
vae_{kk}^{ii}& vae_{kk}^{ij}& vae_{kl}^{ii}& vae_{kl}^{ij}\\
vae_{kk}^{ji}& vae_{kk}^{jj}& vae_{kl}^{ji}& vae_{kl}^{jj}\\
vae_{lk}^{ii}& vae_{lk}^{ij}& vae_{ll}^{ii}& vae_{ll}^{ij}\\
vae_{lk}^{ji}& vae_{lk}^{jj}& vae_{ll}^{ji}& vae_{ll}^{jj}
\end{pmatrix}
\end{split}
\end{align}
\begin{align*}
&v_{c}^{s} = \frac{va_{c}^{s}}{y_{c}^{s}} = 1 - a_{kc}^{is} - a_{kc}^{js} - a_{lc}^{js} - a_{lc}^{is} \hspace{2cm} (c \in k,l \hspace{.5cm} s \in i,j),\\
&\begin{pmatrix}
b_{kk}^{ii}& b_{kk}^{ij}& b_{kl}^{ii}& b_{kl}^{ij}\\
b_{kk}^{ji}& b_{kk}^{jj}& b_{kl}^{ji}& b_{kl}^{jj}\\
b_{lk}^{ii}& b_{lk}^{ij}& b_{ll}^{ii}& b_{ll}^{ij}\\
b_{lk}^{ji}& b_{lk}^{jj}& b_{ll}^{ji}& b_{ll}^{jj}
\end{pmatrix}
=
\begin{pmatrix}
1-a_{kk}^{ii}& -a_{kk}^{ij}& -a_{kl}^{ii}& -a_{kl}^{ij}\\
-a_{kk}^{ji}& 1-a_{kk}^{jj}& -a_{kl}^{ji}& -a_{kl}^{jj}\\
-a_{lk}^{ii}& -a_{lk}^{ij}& 1-a_{ll}^{ii}& -a_{ll}^{ij}\\
-a_{lk}^{ji}& -a_{lk}^{jj}& -a_{ll}^{ji}& 1-a_{ll}^{jj}
\end{pmatrix}^{-1},
\end{align*}
and
\begin{equation*}
a_{cf}^{su} = \frac{inp_{cf}^{su}}{y_{f}^{u}}  \hspace{2cm} (c,f \in k,l \hspace{.5cm} s,u \in i,j),
\end{equation*}
where \(v_{s}^{c}\) gives the share of industry \emph{s}'s value added, \(va_{c}^{s},\) in output, \(y_{s}^{c}\), and \(e_{k}^{i}\) indicates gross exports. Finally, \(a_{su}^{cf}\) denotes the share of inputs, \(inp_{su}^{cf}\), in output.
The elements of the $V(I-A)^{-1}E$ or $vae$ matrix are our estimates for the country-industry level value added
origins of each country-industry's exports.
\textbf{decompr} implements this algorithm into R to automate the process of deriving the matrix. 
Equipped with it, researchers can calculate standard GVC indicators.
Examples include \citet{dahuetal01}'s Vertical Specialisation ratio at the industry-level using the textbf{vertical\_specialisation} function, 
which sums for each country and industry across the value added of all foreign countries and industries,
and \citet{rojoguno12}'s so-called VAX ratio.
Alternatively, the four dimensions of the matrix (source country, source industry, using country, using industry) allow for industry-level gravity-type estimations of value added trade flows.

@@

One interesting application of this decomposition for trade and development uses changes over time in $FVA_FIN$ and $FVA_INT$. 
When low-wage developing countries enter GVCs, they tend to specialize mainly in assembly but try to gradually move up within the value chain.    
To illustrate this, we can reuse the example of the iPhone. 
Most of the value added in the device stems from US design and Japanese technology but it is ultimately assmbled in China.
This means for China that when it enters this GVC, its $FVA_FIN$ starts to increase since it imports a lot of foreign value added, assembles it, and exports a final good: the Iphone.
However, assembly itself does not contain a lot of value added so that the benefit of China initially is small. 
When its technology however improves due to its interaction with Japan and the USA, it might be able to produce actual parts of the phone,
which contain more sophistication, itself and eventually it might even be able to outsource assembly to a cheaper country.
This would imply that it seizes a larger share of the value added, something commonly referred to as upgrading or moving up within the value chain.
In terms of the WWZ decomposition, we would then observe that China's $FVA_FIN$ first goes up and then starts to decline with
a simultaneous increase in $FVA_INT$.

@@

We have designed the package using a modular structure, with an additional user interface function for increased ease of use.
The modular structure enables users to break the computationally intensive analysis process down into several steps.
Furthermore, the modular structure enables users and other developers to build on top of basic data structures which are creaed by the \textbf{load\_tables\_vectors} function when implementing other decompositions or analyses.
However, a wrapper function (\textbf{decomp}) is also provided, which combines the use of the atomic functions into one.
Lastly, in addition to the \textbf{decompr} function, we provide a Rcmdr plugin \parencite{fox2005getting, fox2007extending}, which allows the user to perform the analytical process using a GUI (Grapical User Interface).
All of this, should allow users of the package to adapt the package to their specific needs as the GVC research progresses.
